---
title: "Data Insights Presentation"
output: html_document
date: "2024-10-16"
---

<!-- https://stackoverflow.com/questions/46221816/how-to-integrate-r-projecttemplate-into-r-markdown -->
`r knitr::opts_knit$set(root.dir='..')`

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(knitr)

# https://stackoverflow.com/questions/46221816/how-to-integrate-r-projecttemplate-into-r-markdown
library(ProjectTemplate); load.project()

source("src/01-plots.R", local = knitr::knit_global())

```


```{r notes, include=FALSE}

# NOTES ----

# Our local Joint Forward Plan highlights the core purposes of an integrated care system (ICS). 

# Two of these are:
# - Improving outcomes in population health and care
# - Tackling inequalities in outcomes, experience and access

# The Population Health Management objectives include Cardiovascular Disease (CVD) Prevention – to develop a programme of population health management interventions targeting High Blood Pressure and Cholesterol

# Presentation brief ----

# The joint forward plan can be found here: https://improvinglivesnw.org.uk/norfolk-and-waveney-5-year-joint-forward-plan/

# - Presentation on:
#   - trends in CVD and 
#   - opportunities to reduce variation in primary and secondary prevention.

# You have 10 minutes on the agenda and the deputy director of public health prefers visuals over text

# The slides might include some of the following:

# - trends in CVD outcomes for Norfolk and Waveney 
#   – mortality and hospital admissions 
#     – how does Norfolk and Waveney benchmark with other areas?

# - inequality in CVD outcomes by community across Norfolk and Waveney 
#   – mortality and hospital admissions 
#     – do some areas see more or less than expected than others

# - for local gp practices: 
#     - what does variation/inequality look like?
#     - identify opportunities to improve outcomes in primary prevention – health behaviours e.g. smoking, obesity
#     - identify opportunities to improve outcomes in secondary prevention – detecting and managing high blood pressure, managing patients with high cholesterol and high CVD risk through use of statins

# - is there anything we need to consider that might impact need for CVD related services in future?

# Resources ----

# National:

# - Emergency admissions: 
# https://fingertips.phe.org.uk/search/hospital%20admissions#page/3/gid/1/pat/15/par/E92000001/ati/502/are/E10000020/iid/90988/age/1/sex/4/cat/-1/ctp/-1/yrr/1/cid/4/tbm/1

# Regional:

# - Small area health data: 
# https://www.localhealth.org.uk/
# https://fingertips.phe.org.uk/profile/local-health/data#page/0/gid/1938133185/pat/402/par/E10000020/ati/3/are/E02005537/yrr/1/cid/4/tbm/1



# - National General Practice Profiles: 
# https://fingertips.phe.org.uk/profile/general-practice




# - CVD prevent data and api: 
# https://data.cvdprevent.nhs.uk/data-explorer

# - Health Inequalities:
# https://analytics.phe.gov.uk/apps/health-inequalities-dashboard/

# - Analytical methods, fingertips api and guides etc: 
# https://fingertips.phe.org.uk/profile/guidance/supporting-information/PH-methods

# - Local JSNA reports: 
# https://www.norfolkinsight.org.uk/jsna/document-library/


# Norfolk and Waveney overview ----

# Our Mission
# To help the people of Norfolk and Waveney to live longer, healthier and happier lives.

# Our Goal 1

# To make sure that people can live as healthy a life as possible.

# This means preventing avoidable illness and tackling the root causes of poor health. 
# We know the health and wellbeing of people living in some parts of Norfolk and Waveney is significantly poorer – 
# how healthy you are should not depend on where you live. 
# This is something we must change.

# Our Goal 2

# To make sure that you only have to tell your story once.

# Too often people have to explain to different health and care professionals what has happened in their lives,
# why they need help, the health conditions they have and which medication they are on. 
# Services have to work better together.

# Our Goal 3

# To make Norfolk and Waveney the best place to work in health and care.

# Having the best staff, and supporting them to work well together, will improve the working lives of our staff, 
# and mean people get high quality, personalised and compassionate care.


# Definitions ----

# Cardiovascular disease (CVD): 
# Umbrella name for conditions that affect your heart or circulation. 
# These include high blood pressure, stroke and vascular dementia. 
# Heart disease includes conditions that narrow or block blood vessels (coronary heart disease).


```

## Part 1 - trends in CVD outcomes for Norfolk and Waveney 
## – Mortality and hospital admissions 
## – How does Norfolk and Waveney benchmark with other areas?

---

#### Prevalence

```{r cvd-prev-plots, echo=FALSE, warning=FALSE, message=FALSE}

sex_barplot
sex_age_sd_barplot
age_grp_barplot
dep_quint_barplot
dep_quint_sd_barplot
cat("Source: https://data.cvdprevent.nhs.uk/data-explorer?period=18&level=7&area=8065&indicator=12")

```

---

#### Funnel plot

```{r funnel-plot, echo=FALSE, warning=FALSE, message=FALSE}

funnel_plot
cat("Source: https://fingertips.phe.org.uk/search/hospital%20admissions#page/3/gid/1/pat/15/par/E92000001/ati/502/are/E10000020/iid/90988/age/1/sex/4/cat/-1/ctp/-1/yrr/1/cid/4/tbm/1")

```

#### Region bar plot

```{r region-bar, echo=FALSE, warning=FALSE, message=FALSE}

region_bar
cat("Source: https://fingertips.phe.org.uk/search/hospital%20admissions#page/3/gid/1/pat/15/par/E92000001/ati/502/are/E10000020/iid/90988/age/1/sex/4/cat/-1/ctp/-1/yrr/1/cid/4/tbm/1")

```

```{r echo=FALSE}

```

## Part 2 - Inequality in CVD outcomes by community across Norfolk and Waveney 
## – Mortality and hospital admissions 
## – Do some areas see more or less than expected than others

"Deaths from circulatory disease, all ages, standardised mortality ratio", 
"Deaths from coronary heart disease, all ages, standardised mortality ratio",         
"Deaths from stroke, all ages, standardised mortality ratio"  

Deaths from circulatory disease, coronary heart disease, stroke, all ages, standardised mortality ratio (averaged).

```{r echo=FALSE}

# MSOA = "Middle level Super Output Area"

```


```{r echo=FALSE}

# TODO

# - Add "Recent trend" to part 1 

```

## Part 3 - for GP practices: 
## - What does variation/inequality look like?
## - Identify opportunities to improve outcomes in primary prevention – health behaviours e.g. smoking, obesity
## - Identify opportunities to improve outcomes in secondary prevention – detecting and managing high blood pressure, managing patients with high cholesterol and high CVD risk through use of statins

```{r echo=FALSE}

```

```{r echo=FALSE}

```

```{r echo=FALSE}

```

```{r echo=FALSE}

```

```{r echo=FALSE}

```
